Pleiotropy of gene regulatory variants increases with the number of regulated genes, active cell types and tissues and bound transcription factors

### 2022-11-12 ms/ms03_results.tex
MS
ENH

Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
%
The first criterion was to exclude molecular traits such as proteome or methylome.
%	
The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
%
The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental traits such "employment status" or "self-reported" medical conditions.
%
The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
%
These filters resulted in 413 GWAS (Supplementary table 2).

I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
%
This analysis resulted in 30 261 unique variants and 2,479,065 potential variant colocalizations from 150 GWAS studies (OSF URL).
%
%TODO add PMID to supplementary table with loci_explained_perc
For GWAS studies with more than 10 loci, the explained of loci variate from 0\% to 77\% with an average of 33\% (Supplementary table ST6).
%
Autoimmune diseases such as inflammatory bowel disease, Crohn's disease and Ulcerative colitis showed a colocalization average of 45.4\% in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
%
Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).

@@

I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 2).
%
In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
%
Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).

>>>>

Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
%
The first criterion was to exclude molecular traits such as proteome or methylome.
%	
The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
%
The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental traits such "employment status" or "self-reported" medical conditions.
%
The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
%
These filters resulted in 418 GWAS (Supplementary table 2).
%
I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
%
This analysis tested 139,069 unique variants and ? potential variant/GWAS/eQTL sample and e-gene colocalizations from 418 GWAS studies (OSF URL).
%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc); 139,069
%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc where `PP.H4.abf`>=0.8); 103,551
%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc);
%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc where `PP.H4.abf`>=0.8); 2,477,640
%
I found 103,551 variants or from ? GWAS that colocalized with at least one eQTL.
%select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
%
%TODO add PMID to supplementary table with loci_explained_perc
For GWAS studies with more than 10 loci, the explained of loci variate from 0\% to 77\% with an average of 33\% (Supplementary table ST6).
%
Autoimmune diseases such as inflammatory bowel disease, Crohn's disease and Ulcerative colitis showed a colocalization average of 45.4\% in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
%
Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).

@@

I manually assigned the 418 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 2).

%
In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
%
Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).
